### 2025-02-13 README.Rmd
fix: update renv version
update linux dependencies
update README

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/bbartholdy/mb11CalculusPilot/main?urlpath=rstudio) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825.svg)](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825)

@@

<!--
> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:
-->

> Bartholdy, B.P., Hasselstrøm, J.B., Sørensen, L.K., Casna, M., Hoogland, M., Historisch Genootschap Beemster, & Henry, A.G. (2023). _`r Title`_. Zenodo, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.7649151>

### How to cite

Use the above paper to cite this compendium.

<!--
Please cite this compendium as:

> Bartholdy, B.P., Hasselstrøm, J.B., Sørensen, L.K., Casna, M., Hoogland, M., Historisch Genootschap Beemster, & Henry, A.G., (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.5281/zenodo.7649825>
-->

@@

## How to run in your browser or download and run locally

**If you are running Linux, you will need to install `gfortran`.**

e.g. `sudo apt install gfortran` for Ubuntu.

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself, optionally (but recommended) [RStudio Desktop](https://rstudio.com/products/rstudio/download/), and [Quarto](https://quarto.org/docs/get-started/).

@@

`git clone git@github.com:bbartholdy/mb11CalculusPilot.git`

@@

- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.qmd` and **render** to produce the PDF manuscript (`_output/`),
  or run `quarto::quarto_render("analysis/paper/paper.qmd")` in the R console.

>>>>

[![R-CMD-check](https://github.com/bbartholdy/mb11CalculusPilot/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bbartholdy/mb11CalculusPilot/actions/workflows/R-CMD-check.yaml) [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825.svg)](https://zenodo.org/badge/DOI/10.5281/zenodo.7649825)

@@

> Bartholdy, Bjørn Peare; Hasselstrøm, Jørgen B.; Sørensen, Lambert K.; Casna, Maia; Hoogland, Menno; Historisch Genootschap Beemster; Henry, Amanda G. Multiproxy analysis exploring patterns of diet and disease in dental calculus and skeletal remains from a 19th century Dutch population. Peer Community Journal, Volume 4 (2024), article no. e46. doi : 10.24072/pcjournal.414. https://peercommunityjournal.org/articles/10.24072/pcjournal.414/

@@

## How to reproduce the paper

**Software requirements:**

- [R](https://cran.r-project.org/)
- [Quarto](https://quarto.org/)
- [Git](https://git-scm.com/downloads) (optional)

**Linux dependencies:**

`libx11-dev libicu-dev make zlib1g-dev libcurl4-openssl-dev libssl-dev libfontconfig1-dev libfreetype6-dev libfribidi-dev libharfbuzz-dev libjpeg-dev libpng-dev libtiff-dev pandoc libxml2-dev libgit2-dev`

@@

`git clone https://github.com/bbartholdy/mb11CalculusPilot.git`

@@

- Open the folder in your preferred IDE
- run `renv::restore()` to install the required R packages
- run `quarto::quarto_render("analysis/paper/paper.qmd", output_format = "pdf")` in the R console (requires installing **quarto** R package) or `quarto render analysis/paper/paper.qmd --to pdf` to produce a PDF
- find the pdf in `./analysis/paper/_output`
